Open Questions for Computer Science and Cheminformatics

### 2011-11-21 OpenQuestions.tex
added a hadoop challenge

There exist a multitude of programs generation 3D conformations of molecules. Still, often input formats vary and require
to re-create 3D conformational spaces with other programs. This does not only complicate the maintenance scenario of such
databases, especially with the explosion of public databases, but also creates hugh space and time complexity issues when
searching within such databases. What could be better interfaces, maintenance, data structures, and private/public sharing
scenarios for conformational 3D databases? 
Many software vendors use different solutions for parallizing compute jobs: SGE, PVM, MPI, etc. 
Everyone knowing the enterprise IT approval cycles might know that having more than one parallel processing framework 
is not easy. What could be better ways to streamline parallel processing structures for 
cheminformatics (and molecular modeling) algorithms? Is a cloud really an option? What about SaaS with secured data transfer?
Can this also offer alternative licensing strategies for software suites in this domain?
\end{enumerate}

>>>>

  There exist a multitude of programs generation 3D conformations of
  molecules. Still, often input formats vary and require to re-create
  3D conformational spaces with other programs. This does not only
  complicate the maintenance scenario of such databases, especially
  with the explosion of public databases, but also creates hugh space
  and time complexity issues when searching within such
  databases. What could be better interfaces, maintenance, data
  structures, and private/public sharing scenarios for conformational
  3D databases?

\item \emph{Parallelization of cheminformatics problems \& algorithms}
Many cheminformatics tasks apply algorithms over large input files or
across many molecules. A simple way to parallelize this is to chunk
the input data and let individual threads/node process each chunk. A
trivial solution is to manually chunk the input and submit a series of
jobs to a scheduling system.

Can we employ modern frameworks like Hadoop to perform this? Can we
develop a generalized framework that includes the requisite
cheminformatics tools that allows users to seamless distribute jobs
over Hadoop and other map reduce systems?

Going further, are there cheminformatics problems that can make use of
the map/reduce paradigm at the algorithmic level? More generally, are
cheminformatics algorithms amenable to parallelization (that does not
involve chunking the input data)

\item \emph{Licensing in a parallel world} Many software vendors use
  different solutions for parallizing compute jobs: SGE, PVM, MPI,
  etc.  Is a cloud really an option?  What about SaaS with secured data
  transfer?  Can this also offer alternative licensing strategies for
  software suites in this domain?

\end{enumerate}
